\LaTeX\ Author Guidelines for ICCV Proceedings

### 2019-04-25 chapters/1_intro.tex
Add algorithm notation

Other clustering methods work with so-called \emph{signed graphs}, which include both positive and negative edge weights corresponding to attraction and repulsion between pixels \TODO{REFS}. The advantage of using signed graphs is that balancing attraction and repulsion allows us to perform the clustering without defining additional parameters. This can be done optimally by solving the so-called \emph{multicut optimization problem} or \emph{correlation clustering} [REF?]; however, this problem is NP-hard. For problems of considerable size, good approximations exist \cite{yarkony2012fast,pape2017solving}, and for even larger problems efficient greedy algorithms were recently proposed \cite{levinkov2017comparative,wolf2018mutex}.

>>>>

Other clustering methods work with so-called \emph{signed graphs}, which include both positive and negative edge weights corresponding to attraction and repulsion between pixels \TODO{REFS}. The advantage of using signed graphs is that balancing attraction and repulsion allows us to perform the clustering without defining additional parameters. This can be done optimally by solving the so-called \emph{multicut optimization problem} or \emph{correlation clustering} \cite{kappes2011globally,chopra1991multiway}; however, this problem is NP-hard. For problems of considerable size, good approximations exist \cite{yarkony2012fast,pape2017solving}, and for even larger problems efficient greedy algorithms were recently proposed \cite{levinkov2017comparative,wolf2018mutex}.

### 2019-04-25 chapters/3_algorithm.tex
Add algorithm notation

\begin{minipage}[b]{\textwidth}
        \centering
        \begin{minipage}[b]{0.45\textwidth}
\centering
        \includegraphics[width=\textwidth,trim=0.1in 0.4in 0.2in 0.2in,clip]{./figs/edge_contraction.png} % left bottom right top
\captionof{figure}{\small 
Example of edge contraction: the edge $e_{uv}$ is selected from priority queue; after nodes $u$ and $v$ are merged and $e_{uv}$ is deleted, there are two double edges $e_1$ and $e_2$ left in the graph...  
\label{fig:edge_contraction_and_contr_graph} }
    \end{minipage}\hspace{0.04\textwidth}
    \begin{minipage}[b]{0.45\textwidth}
        \centering
\includegraphics[width=0.8\textwidth,trim=0.in 0.in 0.in 0.in,clip]{./figs/cannot-lin-constraints.png}
\captionof{figure}{\small 
Example of agglomerative clustering on signed graph with and without cannot-link constraints...
\label{fig:algorithm_with_without_CLC}}
    \end{minipage}
\end{minipage}

\section{Unified Graph Agglomerative Clustering}
In this section we introduce the Unified Graph Agglomerative Clustering Algorithm (UGACA) that represents a simple generalized formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive node similarities and signed clustering algorithms using attractive and repulsive cues.}\\
First, we define notation and the concept of cannot-link constraints. Then, we introduce the generalized algorithm and show \UPDATE{how it can be used to introduce several new agglomerative clustering algorithms for graphs with signed edge weights.}

\subsection{Contracted graph and cannot-link constraints} \label{sec:contr_graph_and_CLC}

\begin{itemize}
\item Define graph formalism and clustering $\Pi$. Make comment about special case of unsigned graph
\item Define multicut and associated energy

\item Given a clustering $\Pi$ and a graph $\mathcal{G}(V,E,W)$, the interaction between two clusters $S_1, S_2 \in \Pi$ is usually defined in terms of a \emph{linkage criterion}, i.e. a function of all the edge weights connecting the two clusters:
\begin{equation} \label{eq:linkage_criterion_def}
\begin{gathered}
\mathcal{L}(S_1,S_2) = \mathcal{L}(\mathcal{W})\quad \\
   \text{where} \quad \mathcal{W} = \{ W(e_{uv})| u\in S_1, v\in S_2 \}.
\end{gathered}
\end{equation}

\item A cannot-link-constraint defines a mutual exclusion relationship between two nodes in the graph and is used to specify that the these two nodes should not be associated with the same cluster. (the alg. will introduce them dynamically)
\item Define contracted graph from given graph  $\mathcal{G}$ and clustering $\Pi$; explain edge contraction shown in Fig. 2 and define \emph{update rule} $f(w_1, w_2)$

\end{itemize}

\begin{algorithm}
  \caption{Unified Graph Agglomerative Clustering \TODO{adapt to width}}
   \hspace*{\algorithmicindent} \textbf{Inputs:} graph $\mathcal{G}(V,E,W)$; bool {\color{blue}\texttt{addCannotLink}}  \\
  \hspace*{\algorithmicindent} \textbf{Outputs:} Final clustering and contracted graph $\mathcal{G}'$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]

    % \Procedure{SignedGraphEdgeContr}{{\color{blue}bool \emph{addConstraints}}}
      % \State $\mathcal{G}'\gets \mathcal{G}(V,E^+ \cup E^-)$ \Comment{Initialize the contracted graph}
      \State Initialize contracted graph $\mathcal{G}'$ from $\mathcal{G}$
      \State Initialize \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$
      \State Sort edges in priority queue (PQ) in desc. order of $|W_e|$ 
      \State
      \While{PQ is \textbf{not} empty}
        \State Pop edge $e_{uv}$ in $\mathcal{G}'$ with highest priority $|\tilde{w}|$
        \State
        \If{({\color{ForestGreen}\textbf{$\tilde{w} > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}

          % \State PQ, $\,E_\dagger,\,\, E' \gets$ \textsc{deleteDoubleEdges}($u,v$)

        %   \State Update costs of double edges;
        %   \State Propagate constrained flags of double edges;
          \State Contract $e_{uv}$ in $\mathcal{G}'$
          \For{every new pair of double edges $(e_1,e_2)$}
            \State Get priorities $\tilde{w}_1, \tilde{w}_2$ of $e_1,e_2$
            \State Use update rule $f(\tilde{w}_1,\tilde{w}_2)$ (see Table 1) to 
            \Statex \hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent} update priority of $e_1$
            \State Delete $e_2$ from $\mathcal{G}'$
          \EndFor
          % \State  Replace double edges in $\mathcal{G}'$ with single edges
          % \Statex \hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent} and update priorities
        \EndIf
        \State
        \If{({\color{red}\textbf{$\tilde{w} \leq 0$}}) \textbf{and} {\color{blue}\texttt{addCannotLink}}}
          % \State Flag $e_{uv}$ as MustNotLink
          \State \texttt{canBeMerged}$[e_{uv}] \gets$ \texttt{False}
        \EndIf
      \EndWhile

      % \For{$e  \in E$ in descending order of absolute costs $|W(e)|$}
      %   \If{({\color{ForestGreen}\textbf{$W(e) > 0$}}) \textbf{AND} (edge is not constrained)}
      %     \State Contract edge $e$ in the graph $\mathcal{G}'$;
      %     \State Update costs of double edges;
      %     \State Propagate constrained flags of double edges;
      %     % \For{every new double edge}
      %     %   \State Delete double edges
      %     %   \State Insert new one with updated cost
      %     % \EndFor
      %   \EndIf
      %   \If{({\color{red}\textbf{$W(e) \leq 0$}}) \textbf{AND} ({\color{blue}\emph{addConstraints}})}
      %     \State Flag edge $e$ as constrained
      %   \EndIf
      % \EndFor
      \State
      \State
      \Return $\mathcal{G}'$
      % \State

@@

\subsection{The algorithm}

We now present our first contribution: a unified algorithm for agglomerative graph clustering that can describe in a simple general form many graph partitioning algorithms previously introduced in the literature \UPDATE{and highlights the existence of totally new algorithms}.

We consider a graph $\mathcal{G}(V,E,W)$ with associated signed edge weights, where each node initially represents one cluster.
The proposed graph partitioning algorithm is a general form of agglomerative clustering that outputs a final graph clustering $\Pi$ by progressively merging pairs of neighboring clusters. Intuitively, the algorithm starts by merging clusters with the strongest attractive interaction, corresponding to the highest positive weights in the graph, and it stops when the remaining clusters share only mutual repulsive interactions (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}b} and \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}c}).

A second variant of the proposed algorithm also introduces cannot-link constraints between clusters during the agglomeration and stops when all the remaining clusters have been previously constrained (option \texttt{addCannotLink} in Algorithm \ref{main_alg}).

During the agglomerative process, the interaction between neighboring clusters is defined by the linkage criterion defined in Equation \ref{eq:linkage_criterion_def} and has to be properly updated and recomputed. 
An efficient way of implementing these updates can be achieved by using the presented edge contraction Algorithm \ref{main_alg}, which proceeds as follows.

Given an initial clustering $\Pi$ where each node of the original graph $\mathcal{G}$ represents its own cluster, the algorithm starts by initializing a contracted graph $\mathcal{G}'$ (see Definition \ref{def:contr_graph}) that will represent the clustering $\Pi$ at each iteration of the agglomeration. The algorithm then sorts all edges $e\in E$ by their absolute weight $|W(e)|$ in descending order in a priority queue, so that the most attractive and the most repulsive edges are processed first. It iteratively pops one edge $e_{uv}$ from priority queue and, depending on the value of the priority $\tilde{w}$ associated with $e_{uv}$, does the following: if the edge is attractive and $\tilde{w}>0$, then merge the connected nodes and perform an edge contraction of $e_{uv}$ in $\mathcal{G}'$ (see Sec. \ref{sec:contr_graph_and_CLC} and Fig. \ref{fig:edge_contraction_and_contr_graph}), provided that $e_{uv}$ was not previously flagged as a cannot-link constraint; for every new pair of double edges in $\mathcal{G}'$ update \UPDATE{their priorities} according to one of the update rules listed in Table \ref{tab:linkage-criteria} and update their cannot-link relationships (see example in Fig. \ref{fig:algorithm_with_without_CLC}b from step 3 to 4). If, on the other hand, the edge $e_{uv}$ is repulsive and $\tilde{w}\leq 0$, then flag it as a cannot-link constraint if and only if the algorithm input-option \texttt{addCannotLink} is true and mutual exclusion relationships should be enforced.

In the Supplementary material we provide a more detailed description of the algorithm (Sec. \ref{sec:detailed_algorithm}) and we also comment on its computational complexity (Sec. \ref{sec:complexity}).

 In the special case of an unsigned graph %$\mathcal{G}(V,E,W:E\rightarrow \mathbb{R}^+)$ 
 presenting strictly positive similarities between nodes, the algorithm returns only one single cluster as an output (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}a}), \UPDATE{together with a merging tree representing the merging order and defining a hierarchy of clusters} \TODO{use def of merging process of contracted graph to define merging tree?}.

@@

@@

            & & &  \thead{\textsc{No}} & \thead{\textsc{Yes}} \\        
            % \cmidrule[0.3em]{3-5}
            % \midrule[0.15em]
            \midrule\midrule
            Sum: & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \tilde{w}_1+\tilde{w}_2$} & \thead{Sum Linkage\\Hierarchical Clust.} & \thead{Greedy Additive \\ Edge Contraction \\\cite{levinkov2017comparative}} & \thead{Greedy \\Fixation \\\cite{levinkov2017comparative}} \\ \midrule

            \makecell[r]{Abs. \\ max:} & \thead[l]{
            $
            f(\tilde{w}_1,\tilde{w}_2) = \begin{cases} 
            \tilde{w}_1 & \text{if}\,\, |\tilde{w}_1|>|\tilde{w}_2|\\
            \tilde{w}_2 & \text{otherwise}
             \end{cases} 
            $}
               & \thead{Single Linkage\\Hierarchical Clust.\\ \cite{lance1967general}} & \thead{Mutex \\Watershed \\\cite{wolf2018mutex}} & \thead{Mutex \\Watershed \\\cite{wolf2018mutex}} \\ \midrule
            \makecell[r]{Mean:} & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \mathrm{wAvg}\{ \tilde{w}_1, \tilde{w}_2 \} $}                                 & \thead{ Average Linkage\\ Hierarchical Clust. \\(UPGMA) \cite{lance1967general}} & \thead{Average Linkage\\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}\\ \midrule

            Max: & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \max \{ \tilde{w}_1, \tilde{w}_2 \}  $}                                 & \thead{Single Linkage\\Hierarchical Clust.\\ \cite{lance1967general}} & \thead{Single Linkage \\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}\\ \midrule

            Min:& \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \min \{ \tilde{w}_1, \tilde{w}_2 \}  $}                                 & \thead{Complete Linkage\\ Hierarchical Clust. \\ \cite{lance1967general}}  & \thead{Complete Linkage \\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}

>>>>

\begin{minipage}[t]{\textwidth}
        \centering
        \begin{minipage}[t]{0.45\textwidth}
\centering
        \includegraphics[width=\textwidth,trim=0.1in 0.4in 0.2in 0.2in,clip]{./figs/edge_contraction.png} % left bottom right top
\captionof{figure}{{\small 
Example of edge contraction. The first row represents the original graph $\mathcal{G}$, the clustering $\Pi$ (purple lines) and the edges not on the cut (highlighted in green) before and after the contraction. The second row shows the corresponding contracted graph $\mathcal{G'}$ and cannot-link constraints (edges with red bars). In step (b), edge $e_{uv}$ is contracted: nodes $u$ and $v$ are merged and $e_{uv}$ is deleted from $\mathcal{G'}$. In (c), the two double edges $e_1$ and $e_2$ left in $\mathcal{G'}$ after the contraction are replaced by a single constrained edge with updated cost $f(c_1, c_2)$, where $f$ is one of the update rules defined in Table \ref{tab:linkage-criteria}.} 
\label{fig:edge_contraction_and_contr_graph} }
    \end{minipage}\hspace{0.04\textwidth}
    \begin{minipage}[t]{0.45\textwidth}
        \centering
\includegraphics[width=0.8\textwidth,trim=0.in 0.in 0.in 0.in,clip]{./figs/cannot-lin-constraints.jpg}
\captionof{figure}{\small 
Example of agglomerative clustering on signed graph with and without cannot-link constraints...
\label{fig:algorithm_with_without_CLC}}
    \end{minipage}
\end{minipage}

\section{Generalized framework for Agglomerative Clustering of Signed Graphs}
In this section we introduce the Unified Graph Agglomerative Clustering Algorithm (UGACA) that represents a simple \UPDATE{generalized} formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive node similarities and signed clustering algorithms using attractive and repulsive cues.}\\
First, we define notation and the concept of cannot-link constraints. Then, we introduce the generalized algorithm and show \UPDATE{how it can be used to introduce several new agglomerative clustering algorithms for graphs with signed edge weights.}

\subsection{Notation and graph formalism} \label{sec:notation}

Given a simple graph $\mathcal{G}(V,E)$, we call the set $\Pi$ a \emph{clustering} with $K$ clusters if $V = \cup_{i=1}^K S_i $, $S_i \cap S_j = \emptyset$ for every $i\neq j$ and every $S \in \Pi$ induces a connected subgraph of $\mathcal{G}$.

In this work, we consider the problem of clustering a weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. The weight function $w^+: E \rightarrow \mathbb{R}^+$ associates a positive scalar attribute $w_e^+\in \mathbb{R}^+$ to each edge $e \in E$ representing a merge affinity or a similarity measure: the higher this number, the higher the inclination of the two incident vertices to be assigned to the same cluster\footnote{Note that several formalisms defining positively weighted graphs associate to each edge a positive scalar that represents a metric or distance on the graph. In these formalisms, the \emph{lower} the edge weight, the higher the attraction between the two linked nodes, contrary to our definition of $w^+: E \rightarrow \mathbb{R}^+$.}. On the other hand, $w^-: E \rightarrow \mathbb{R}^+$ associates a split tendency $w_e^- \in \mathbb{R}^+$ to each edge: the higher this weight, the greater the desire of the incident vertices to be in different clusters. 
Often, graphs with both attractive and repulsive edge attributes are described in terms of a \emph{signed} cost function $\cost: E \rightarrow \mathbb{R}$ that combines both edge weights into one signed scalar number: 
\begin{equation}
\cost_e = w_e^+ - w_e^- \in \mathbb{R}, \quad \quad \forall e \in E
\end{equation}
Graphs that have only positive weights can be clearly seen as special cases of the previously defined graph $\mathcal{G}(V,E,w^+,w^-)$, when $w^-_e=0$ and $c_e \geq 0$ for every $e \in E$.

\paragraph{Linkage criterion}\label{par:linkage_criterion_def} We call two clusters $S_1,S_2$ \emph{adjacent} if there exists at least one edge $e_{uv}\in E$ in the original graph connecting a node $u\in S_1$ to a node $v\in S_2$:
\begin{equation}
 \text{$S_1, S_2 \in \Pi$ are \emph{adjacent}} \quad \Longleftrightarrow  \quad  \mathcal{B}(S_1,S_2) \equiv \{ e_{uv} \in E | u \in S_1, \, v \in S_2 \} \neq \emptyset
\end{equation}
where we defined the set of edges connecting two adjacent clusters as \UPDATE{\emph{boundary}} \UPDATE{$\mathcal{B}(S_1,S_2)$}. The interaction between two clusters $S_1, S_2$ is usually defined as a function, named \emph{linkage criterion}, depending on the weights of all edges in the boundary $\mathcal{B}(S_1,S_2)$. In hierarchical agglomerative clustering and graphs with only positive weights, some examples of linkage criterion statistics are \emph{maximum} (single-linkage clustering), \emph{minimum} (complete-linkage clustering) and arithmetical average (average-linkage clustering).

\paragraph{Multicut and correlation clustering} For any clustering $\Pi$ of $\mathcal{G}$, we define as $E^0_\Pi$ the set of edges linking nodes in the same cluster, and as $E_\Pi^1$ the complementary set of edges whose linked nodes belong to distinct clusters:
\begin{equation}
E_\Pi^0 \equiv \{ e_{uv} \in E \,|\, \exists S \in \Pi : u \in S \, \text{and} \, v \in S \}, \qquad E^1_\Pi \equiv E \setminus E^0_\Pi.
\end{equation}
% \begin{align}
% E_\Pi^0 &= \{ e_{uv} \in E \,|\, \exists S \in \Pi : u \in S \, \text{and} \, v \in S \}, \\
% E^1_\Pi &= E \setminus E^0_\Pi.
% \end{align}
The set of edges $E_\Pi^1$ is known as the \emph{multicut} of $\mathcal{G}$ associated to the clustering $\Pi$. The instance of the NP-hard \emph{weighted correlation clustering} or \emph{minimum cost multicut problem} w.r.t. $\mathcal{G}(V,E,w^+, w^-)$ \cite{kappes2011globally,chopra1991multiway,andres2015lifting} is the task of finding a clustering that optimally balance the attraction and repulsion in the graph and is given by the following optimization problem/binary integer program:
\begin{equation}
% \min_\Pi \texttt{MC}(\Pi) \equiv
 \min_\Pi \sum_{e\in E} \cost_e x_e^\Pi,  \qquad \text{where} \quad x^\Pi_e = 
 \begin{cases} 
 1 & \text{if } e\in E^1_\Pi \\
 0 & \text{otherwise}.
 \end{cases}
\end{equation}
In the next chapters we will use this objective as a way to measure how balanced is a clustering found by the tested agglomerative clustering algorithms.

\begin{algorithm}[t]
  \caption{Generalized Agglomerative Clustering Algorithm for Signed Graphs}
   \hspace*{\algorithmicindent} \textbf{Inputs:} graph $\mathcal{G}(V,E,W)$; boolean {\color{blue}\texttt{addCannotLink}}  \\
  \hspace*{\algorithmicindent} \textbf{Outputs:} Final clustering $\Pi$ and contracted graph $\mathcal{G}'$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
    % \Procedure{SignedGraphEdgeContr}{{\color{blue}bool \emph{addConstraints}}}
      % \State $\mathcal{G}'\gets \mathcal{G}(V,E^+ \cup E^-)$ \Comment{Initialize the contracted graph}
      \State Initialize clustering $\Pi=\{\{v_1\}, \ldots, \{v_N\}\}$ and contracted graph $\mathcal{G}' \gets \mathcal{G}$
      \State Initialize \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$
      \State Sort edges in priority queue (PQ) in descending order of absolute costs $|w^+_e - w^-_e|$ 
      \State
      \While{PQ is \textbf{not} empty}
        \State Pop edge $e_{uv}$ in $\mathcal{G}'$ with highest absolute cost $|\cost|$
        \If{({\color{ForestGreen}\textbf{$\cost > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}
          % \State PQ, $\,E_\dagger,\,\, E' \gets$ \textsc{deleteDoubleEdges}($u,v$)
        %   \State Update costs of double edges;
        %   \State Propagate constrained flags of double edges;
          \State Contract $e_{uv}$ in $\mathcal{G}'$ and merge clusters in $\Pi$
          \For{every new pair of double edges $(e_1,e_2)$}
            % \State Get costs $\cost_1, \cost_2$ of $e_1,e_2$
            \State Delete $e_2$ from $\mathcal{G}'$ and update cost of $e_1$ with update rule $f(\cost_{e_1},\cost_{e_2})$ def. in Tab. 1 
            % \Statex \hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent} update priority of $e_1$
            % \State Delete $e_2$ from $\mathcal{G}'$
          \EndFor
          % \State  Replace double edges in $\mathcal{G}'$ with single edges
          % \Statex \hspace{\algorithmicindent}\hspace{\algorithmicindent}\hspace{\algorithmicindent} and update priorities
        \EndIf
        % \State
        \If{({\color{red}\textbf{$\cost \leq 0$}}) \textbf{and} {\color{blue}\texttt{addCannotLink}}}
          % \State Flag $e_{uv}$ as MustNotLink
          \State \texttt{canBeMerged}$[e_{uv}] \gets$ \texttt{False}
        \EndIf
      \EndWhile
      % \For{$e  \in E$ in descending order of absolute costs $|W(e)|$}
      %   \If{({\color{ForestGreen}\textbf{$W(e) > 0$}}) \textbf{AND} (edge is not constrained)}
      %     \State Contract edge $e$ in the graph $\mathcal{G}'$;
      %     \State Update costs of double edges;
      %     \State Propagate constrained flags of double edges;
      %     % \For{every new double edge}
      %     %   \State Delete double edges
      %     %   \State Insert new one with updated cost
      %     % \EndFor
      %   \EndIf
      %   \If{({\color{red}\textbf{$W(e) \leq 0$}}) \textbf{AND} ({\color{blue}\emph{addConstraints}})}
      %     \State Flag edge $e$ as constrained
      %   \EndIf
      % \EndFor
      % \State
      \State
      \Return $\Pi$, $\mathcal{G}'$
      % \State

@@

\subsection{Generalized agglomerative clustering framework} \label{sec:algorithm}

We now present our main contribution: a generalized framework for agglomerative graph clustering that can describe in a simple form several graph partitioning algorithms previously introduced in the literature and highlights the existence of new ones.

Given a graph $\mathcal{G}(V,E,w^+, w^-)$, the proposed generalized clustering uses a bottom-up approach starting with each node assigned to its own cluster and iteratively merging pairs of adjacent clusters. Intuitively, the algorithm starts by merging clusters with the strongest attractive interaction, corresponding to the highest positive weights in the graph, and it stops when the remaining clusters share only mutual repulsive interactions (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}b} and \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}c}).

A second variant of the proposed algorithm also introduces \emph{cannot-link constraints}, representing mutual exclusion relationships between pairs of nodes that cannot be associated with the same cluster in the final clustering. This variant (Algorithm \ref{main_alg} with \texttt{addCannotLink=True}) stops when all the remaining clusters have been previously constrained.

% The algorithm that we will present in the next section iteratively performs a sequence of so-called \emph{edge contractions} on the original graph $\mathcal{G}(V,E,w^+, w^-)$.

% The graph clustering algorithm we will present in Sec. \ref{sec:algorithm} is a bottom-up approach starting with each node assigned to its own cluster and iteratively merging clusters. During the agglomeration, the current clustering $\Pi$ is represented by a \emph{contracted graph} $\mathcal{G'}(\mathcal{G}, \Pi)$, such that each of its nodes represents a cluster $S \in \Pi$ and \emph{adjacent} clusters are linked by an edge.

\paragraph{Edge contraction and update rules} During the agglomerative process, the interaction between adjacent clusters has to be properly updated and recomputed.  % given by the linkage criterion defined in Sec. \ref{sec:notation} and
An efficient way of implementing these updates can be achieved by representing the clustering $\Pi$ with a \emph{contracted graph} $\mathcal{G'}(\mathcal{G}, \Pi)$, such that each of its nodes is associated to a cluster $S \in \Pi$, edges represent adjacency-relationships between clusters and edge weights inter-cluster interactions.
\UPDATE{Every time a pair of neighboring clusters is merged, the contracted graph is updated by performing an edge contraction and merging two nodes. After the edge contraction, the only cluster interactions that need to be updated are those between nodes connected by double edges in the contracted graph (see Fig. \ref{fig:edge_contraction_and_contr_graph}, step b). For simple linkage criteria, these updates can be easily computed with an update rule $f$ that depends only on the interactions at the previous step }
% As a result, the merged node could be linked to some of its neighbors by double edges . , representing the fact that their interaction should be updated. This update can be easily achieved by using an \emph{update rule} depending only on the weights of the double edges. 
\TODO{such a mess...}
\UPDATE{In this way we do not need to recompute the interaction by using the linkage criterion over all the edges of the original graph, that would be computationally expensive.}

\paragraph{The algorithm} The proposed Algorithm \ref{main_alg} proceeds as follows. Given an initial clustering $\Pi$ where each node of the original graph $\mathcal{G}$ represents its own cluster  \TODO{repetition}, the algorithm starts by initializing a contracted graph $\mathcal{G}'$ (see Definition \ref{def:contr_graph}) that will represent the clustering $\Pi$ at each iteration of the agglomeration. The algorithm then sorts all edges $e\in E$ by their absolute weight $|W(e)|$ in descending order in a priority queue, so that the most attractive and the most repulsive edges are processed first. It iteratively pops one edge $e_{uv}$ from priority queue and, depending on the value of the priority $\tilde{w}$ associated with $e_{uv}$, does the following: if the edge is attractive and $\tilde{w}>0$, then merge the connected nodes and perform an edge contraction of $e_{uv}$ in $\mathcal{G}'$ (see Sec. \ref{sec:notation} and Fig. \ref{fig:edge_contraction_and_contr_graph}), provided that $e_{uv}$ was not previously flagged as a cannot-link constraint; for every new pair of double edges in $\mathcal{G}'$ update \UPDATE{their priorities} according to one of the update rules listed in Table \ref{tab:linkage-criteria} and update their cannot-link relationships (see example in Fig. \ref{fig:algorithm_with_without_CLC}b from step 3 to 4). If, on the other hand, the edge $e_{uv}$ is repulsive and $\tilde{w}\leq 0$, then flag it as a cannot-link constraint if and only if the algorithm input-option \texttt{addCannotLink} is true and mutual exclusion relationships should be enforced.
In the Supplementary material we provide a more detailed description of the algorithm (Sec. \ref{sec:detailed_algorithm}) and we also comment on its computational complexity (Sec. \ref{sec:complexity}).

 In the special case of a graph with only positive edge interactions, i.e. $w_e^-=0$ $\forall e\in E$, %$\mathcal{G}(V,E,W:E\rightarrow \mathbb{R}^+)$ 
 the algorithm performs a standard agglomerative hierarchical clustering by returning only one single cluster (see Fig. \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}a}) and a hierarchy of clusters defined by the order in which the clusters were merged. 
 %\TODO{use def of merging process of contracted graph to define merging tree?}.

@@

% Given a clustering $\Pi$ and a graph $\mathcal{G}(V,E,W)$, the interaction between two clusters $S_1, S_2 \in \Pi$ is usually defined in terms of a \emph{linkage criterion}, i.e. a function of all the edge weights connecting the two clusters:
% \begin{equation} 
% \begin{gathered}
% \mathcal{L}(S_1,S_2) = \mathcal{L}(\mathcal{W})\quad \\
%    \text{where} \quad \mathcal{W} = \{ W(e_{uv})| u\in S_1, v\in S_2 \}.
% \end{gathered}
% \end{equation}

@@

            & & &  \thead{\textsc{No}} & \thead{\textsc{Yes}} \\        
            % \cmidrule[0.3em]{3-5}
            % \midrule[0.15em]
            \midrule\midrule
            Sum: & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \tilde{w}_1+\tilde{w}_2$} & \thead{Sum Linkage\\Hierarchical Aggl.\\Clustering} & \thead{Greedy Additive \\ Edge Contraction \\\cite{levinkov2017comparative}} & \thead{Greedy \\Fixation \\\cite{levinkov2017comparative}} \\ \midrule

            \makecell[r]{Abs. \\ max:} & \thead[l]{
            $
            f(\tilde{w}_1,\tilde{w}_2) = \begin{cases} 
            \tilde{w}_1 & \text{if}\,\, |\tilde{w}_1|>|\tilde{w}_2|\\
            \tilde{w}_2 & \text{otherwise}
             \end{cases} 
            $}
               & \thead{Single Linkage\\Hierarchical Aggl.\\Clustering \cite{lance1967general}} & \thead{Mutex \\Watershed \\\cite{wolf2018mutex}} & \thead{Mutex \\Watershed \\\cite{wolf2018mutex}} \\ \midrule
            \makecell[r]{Mean:} & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \mathrm{weightAvg}\{ \tilde{w}_1, \tilde{w}_2 \} $}                                 & \thead{ Average Linkage\\ Hierarchical Clust. \\(UPGMA) \cite{lance1967general}} & \thead{Average Linkage\\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}\\ \midrule

            Max: & \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \max \{ \tilde{w}_1, \tilde{w}_2 \}  $}                                 & \thead{Single Linkage\\Hierarchical Aggl.\\Clustering \cite{lance1967general}} & \thead{Single Linkage \\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}\\ \midrule

            Min:& \thead[l]{$f(\tilde{w}_1,\tilde{w}_2) = \min \{ \tilde{w}_1, \tilde{w}_2 \}  $}                                 & \thead{Complete Linkage\\ Hierarchical Aggl.\\Clustering \cite{lance1967general}}  & \thead{Complete Linkage \\Signed Aggl. Clust. \\ (\textbf{NEW})} & \thead{\textbf{NEW}}
